A thesis draft

### 2018-10-14 draft.tex
Add content about 5G

\usepackage{todonotes}

@@

 % This is just a draft, nothing suitable for the final product: sources are in
 % form of a footnote

@@

 Through the history of internet, many factor changed and new innovations came,
 but at the end of the day a foundamental key remained: TCP/IP performance are,
 indeed, what the users unconsciously want.\todo{Idea: talk about how TCP/IP
   performes quite well in wired connections, while in wireless one, such as
   wireless/satellite, sucks ass} In this context network optimization plays an
 important role, and it is for this reason that network packets are processed
 while traversing the network in order to make their transmission as fast as
 possible. On top of that, mobile traffic is continuously surging every year,
 with the backbones having more and more data to process. In this context,
 in-hardware VNF solutions are becoming less efficent. Their time-to-deploy, in
 fact, is very high compared to the requirements of the network. With this
 trend, in-hardware solutions will not be able to keep up. Is with this
 considerations that software solutions are being proposed to become the new way
 to deploy network components. While less efficent, they are faster to bring
 them up, offering an unprecedented flexibility\todo{CITE: SDN/NFV-Based Mobile
   Packet Core Network Architectures: A Survey}. Furthermore, new virtualization
 technologies such as Cloud computing and Docker are becoming more and more
 mature for being employed in large-scale production environments, where new
 frameworks, called orchestrators (e.g. Kubernetes, Docker Swarm), allow to
 easily manage the application lifecyle, i.e. deployment, scale up/down and
 finally, removal.

>>>>

\usepackage{todonotes}
\usepackage{graphicx}

@@

 % This is just a draft, nothing suitable for the final product: sources are in
 % form of a todo

@@

 Through the history of internet, many factors changed and new innovations came,
 but at the end of the day a foundamental key remained: connection performances
 are, indeed, what the users really want.\todo{Idea: talk about how TCP/IP
   performes quite well in wired connections, while in wireless one, such as
   wireless/satellite, sucks ass} In this context network optimization plays an
 important role, and it is for this reason that network packets are elaborated,
 compressed and processed while traversing the network in order to make their
 transmission as fast as possible. On top of that, mobile traffic is
 continuously surging every year, with the backbones having more and more data
 to deliver, making in-hardware backbone solutions less efficent. Their
 time-to-deploy, in fact, is very high compared to the requirements of the
 network. With this trend, in-hardware solutions will not be able to keep up. Is
 with this considerations that software a software approach is being proposed to
 become the new way to deploy network components. While less efficent, they are
 faster to bring them up, offering an unprecedented flexibility\todo{CITE:
   SDN/NFV-Based Mobile Packet Core Network Architectures: A Survey}.
 Furthermore, new virtualization technologies such as Cloud computing and Docker
 are becoming more and more mature for being employed in large-scale production
 environments, where new frameworks, called orchestrators (e.g. Kubernetes,
 Docker Swarm), allow to easily manage the application lifecyle, i.e.
 deployment, scale up/down and finally, removal.

\section{The upcoming connectivity standard: 5G}
The continuous innovation in the mobile network connectivity is leading to the
creation of a new standard, the 5G, which is estimated to arrive to the market
constumer in 2020\todo{CITE}. Lead by the Next Generation Mobile Network (NGMN)
alliace, a group composed by the major players in the field of mobile
connectivity, the 5G aims to offer at the end-user a new way to browse the web,
download and watch interactive content. Relatively to 4G, 5G points to have data
rates $10$ times better, with $10$ times smaller end-to-end latency and
increased connection densisty by $100$ times\todo{CITE}.

\begin{table}[t]
\centering
\resizebox{\textwidth}{!}{%
  \begin{tabular}{p{4cm}|p{5,5cm}|p{5,5cm}}
\textbf{Attribute}                                                    & \textbf{LTE capability}                                                                                                                                     & \textbf{Improvement needed to meet NGMN requirements}                 \\ \hline
\textbf{Data rate (per user)}                                         & Up to 100 Mb/s on average Peaks of 600 Mb/s (Cat 11/12).                                                                                                    & 10X expected on average and peak rates and 100X expected on cell edge \\
\textbf{End-to-end latency}                                           & 10 ms for two-way RAN (pre-scheduled). Typically, up to 50 ms end-to-end if other factors are considered (e.g., transmission, CN, internet, proxy servers). & 10X (smaller)                                                         \\
\textbf{Mobility}                                                     & Functional up to 350 km/h (for certain bands up to 500 km/h). No support for civil aviation.                                                                & 1.5X                                                                  \\
\textbf{\begin{tabular}[c]{@{}l@{}}Connection\\ density\end{tabular}} & Typically $\sim$2,000 active users/km2.                                                                                                                     & 100X                                                                 
\end{tabular}%
}
\caption[5G improvement over LTE]{An extract from the official 5G white paper illustating the improvements of 5G relatively to LTE connections.}
\label{chap:intro:table:ltevs5g}
\end{table}

Part of the new requirements can be satisfied using a large radio spectrum with
higher frequencies. But the utilization of higher frequencies mean that the
radio signals can be easly interrupted by physical objects, like buildings and
many geographical elements (such as hills and mountains), clashing with the
expectation of an ever-reachable connectivity. It is here where virtualization
plays an important role. In fact, the re-design of some network components today
existing via hardware can transform a monolitic networking approach to a modular
one, exploting the flexibility that Virtual Network Functions (VNFs)\todo{CITE
  VNF-P: A Model for Efficient Placement of Virtualized Network Functions} can
offer thanks to virtualization, closing the gap to the use-case fullfilment
defined by the NGMS alliance\todo{CITE: NGMN View on 5G architecture}.

\subsection{5G architecture}

With the constraints placed on the requirements formulated by the NGMN alliance,
5G envisage a multi-layered architecture, based on three main layer:\todo{CITE:
  NGMN View on 5G architecture}
\begin{itemize}
\item \textbf{infrastructure resouce layer}: physical resources that are exposed
  via a virtualized interface, and that can be monitored using specific APIs
\item \textbf{business enablement layer}: where a library of functions and
  deployment is contained, and its configuration is accessible via APIs
\item \textbf{business application layer}: layer that contains specific
  applications and services of the operator
\end{itemize}

\subsubsection{Network slicing}
The role of a ``network slice'' in a 5G architecture is to specificatly handle a
Control-place of a praticular service (e.g. smartphones traffic, autonomous
driving, massive IoT), deploying resources in a manner that assure the required
latency, security and reliability. While some legacy of very peculiar services
could require specific hardware, the common resources between services could be
shared in a virtualized way, providing auto-scaling capabilities in services
that are under heavy network pressure.
